State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-04-14 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
trying to solve the conflict
working on estimation section
general updates to paper

 The accuracy of the approximation is linked to the time step used in the temporal discretization of the system, where the state transition appears more linear with a finer resolution. To explain the operation of the UKF we begin by deriving the traditional Kalman filter with the state-space model
\begin{eqnarray}
	\mathbf{x}_t = q(\mathbf{x}_t)\boldsymbol{\theta} - \xi\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{eqnarray}
where $\boldsymbol{\theta}$ and $\xi$ are the systems parameters. In forwards iteration since the observation equation is linear in parameters we only need to use Unsented transformation in prediction step. The update step equations remains the same as the standard Kalman filtering formulations. In backward smoothing path, estimation from the forward filtering is used to form the new set of sigma points in a similar way to the filterin step, however the predicted cross-covariance of the states need to be calculated. A summary of the algorithm is given in Algorithm 
\begin{algorithm}
\caption{The Unscented Kalman Filter}\label{CPTEMDAlgorithm}
\begin{algorithmic}[1]
\State Initialise
\begin{equation*}
\hat{\mathbf x}_0, \mathbf P_0
\end{equation*}
\State Forward iteration
For $t \in \left\lbrace 0,\cdots, N\right\rbrace $,
\begin{equation*}
\mathcal X_t^f=[\hat{\mathbf x}_t^f \quad \hat{\mathbf x}_t^f+\gamma\sqrt{\mathbf P_t^f} \quad \hat{\mathbf x}_t^{f}-\gamma\sqrt{\mathbf P_t^f}]
\end{equation*}
Propogate the sigma points through the state equation
\begin{equation}
\mathcal X_{i,t+1|t}^{f*}=Q(\mathcal X_t^f,u_t)
\end{equation}
\begin{equation*}
 \hat{\mathbf x}_{k +t}^{f-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1|t}^{f*}
\end{equation*}
\begin{equation*}
 \mathbf P_{t +1}^-=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e
\end{equation*}
\begin{equation*}
\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_v)^{-1}
\end{equation*}
\begin{equation*}
\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-})
\end{equation*}
\begin{equation*}
 \mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-}
\end{equation*}
\State Backward iteration
\State initialisation
\begin{equation}
 \mathbf P_T^s= \mathbf P_T^f
\end{equation}
\begin{equation}
\hat{\mathbf x}^s_T= \hat{\mathbf x}^f_T
\end{equation}
For $t \in \left\lbrace T-1, \cdots, 0 \right\rbrace $
\begin{equation*}
\mathcal X_t^s=[\hat{\mathbf x}_t^f \quad \hat{\mathbf x}_t^f+\gamma\sqrt{\mathbf P_t^f} \quad \hat{\mathbf x}_t^f-\gamma\sqrt{\mathbf P_t^f}]
\end{equation*}
\State Propogate the sigma points through the state equation similar to the forward step and compute

\end{algorithmic}
\end{algorithm}
\section{Results}

@@

\subsection{proof}

@@

&=&\mathbf{\Gamma}^{-1}\int\int\limits_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top} 

@@

\begin{tabular} {c|c} 
	\hline\hline
	Symbol & Description \\ \hline
	$v(\mathbf{r},t)$ & mean membrane potential \\
	$g(\mathbf{r},t)$ & average action potential rate \\
	$\mathbf{r}$ & spatial location \\
	$t$ & time (s) \\
	$h(t)$ & post-synaptic response kernel \\
	$\eta(t)$ & Heaviside function \\
	$\zeta$ & inverse synaptic time constant \\
	$w(\mathbf{r},\mathbf{r}')$ & spatial connectivity kernel \\
	$f(\mathbf{r},t)$ & firing function rate \\
	$u(\mathbf{r},t)$ & external input \\
	$\Omega$ & spatial domain \\
	$f_{max}$ & maximal firing rate \\
	$\varsigma$ & slope of sigmoidal activation function \\
	$v_0$ & firing threshold \\
	$\delta(t)$ & Dirac-delta function \\
	$L$ & temporal differential operator \\
	$e(\mathbf{r},t)$ & field disturbance, covariance $\gamma$\\
	$m(\mathbf{r},\mathbf{r}')$ & sensor kernel, variance $\sigma_m^2$ \\
	$\epsilon(\mathbf{r}_n,t)$ & observation noise, covariance $\Sigma_\epsilon$ \\
	$y(\mathbf{r}_n,t)$ & obseration \\
	$n$ & sensor index $n=1,..,N$ \\
	$T$ & time step \\
	$\xi$ & time constant parameter \\
	$\mathbf{\phi(r)}$ & vector of Gaussian basis functions \\
	$\mathbf{x}_t$ & state vector at time $t$ \\
	$\mathbf{\psi}$ & vector of connectivity kernel basis functions \\
	$\theta$ & vector of connectivity kernel parameters \\
	$\Gamma$ & inner product of field basis functions \\
	$q()$ & state function \\
	$k()$ & maps into to discrete state of input function \\
	$\mathbf{e}_t$ & state disturbance, covariance $\Sigma_e$ \\
	$\mathbf{C}$ & observation matrix \\
	$V(\nu)$ & Spectrum of the dynamic field \\ 
	$\nu$ & Spatial frequency \\
	$\nu_c$ & Spatial cut-off frequency \\
\end{tabular}

>>>>

 The accuracy of the approximation is linked to the time step used in the temporal discretization of the system, where the state transition appears more linear with a finer resolution. To explain the operation of the UKF we begin by deriving the traditional Kalman filter with the state-space model
\begin{eqnarray}
	\mathbf{x}_t = q(\mathbf{x}_t)\boldsymbol{\theta} - \xi\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{eqnarray}
where $\boldsymbol{\theta}$ and $\xi$ are the systems parameters. In forwards iteration since the observation equation is linear in parameters we only need to use Unsented transformation in prediction step. The update step equations remains the same as the standard Kalman filtering formulations. In backward smoothing path, estimation from the forward filtering is used to form the new set of sigma points in a similar way to the filterin step, however the predicted cross-covariance of the states need to be calculated. A summary of the algorithm is given in Algorithm 
\begin{algorithm}
\caption{The Unscented Kalman Filter}\label{CPTEMDAlgorithm}
\begin{algorithmic}[1]
\State Initialise
\begin{equation*}
\hat{\mathbf x}_0, \mathbf P_0
\end{equation*}
\State Forward iteration
For $t \in \left\lbrace 0,\cdots, N\right\rbrace $,
\begin{equation*}
\mathcal X_t^f=[\hat{\mathbf x}_t^f \quad \hat{\mathbf x}_t^f+\gamma\sqrt{\mathbf P_t^f} \quad \hat{\mathbf x}_t^{f}-\gamma\sqrt{\mathbf P_t^f}]
\end{equation*}
Propogate the sigma points through the state equation
\begin{equation}
\mathcal X_{i,t+1|t}^{f*}=Q(\mathcal X_t^f,u_t)
\end{equation}
\begin{equation*}
 \hat{\mathbf x}_{t+1}^{f-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1|t}^{f*}
\end{equation*}
\begin{equation*}
 \mathbf P_{t +1}^{f-}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e
\end{equation*}
\begin{equation*}
\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_v)^{-1}
\end{equation*}
\begin{equation*}
\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-})
\end{equation*}
\begin{equation*}
 \mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-}
\end{equation*}
\State Backward iteration
\State initialisation
\begin{equation}
 \mathbf P_T^s= \mathbf P_T^f
\end{equation}
\begin{equation}
\hat{\mathbf x}^s_T= \hat{\mathbf x}^f_T
\end{equation}
For $t \in \left\lbrace T-1, \cdots, 0 \right\rbrace $
\begin{equation*}
\mathcal X_t^s=[\hat{\mathbf x}_t^f \quad \hat{\mathbf x}_t^f+\gamma\sqrt{\mathbf P_t^f} \quad \hat{\mathbf x}_t^f-\gamma\sqrt{\mathbf P_t^f}]
\end{equation*}
\State Calculate a new set of  sigma points, $\mathcal X_{i,t+1|t}^{b*}$ similar to the forward step and compute

\end{algorithmic}
\end{algorithm}
\section{Results}
We simulated a neural field to the generate data used in the state and parameter estimation. An example of the simulated field can be seen in Figure~\ref{}. The Figure demonstrates the experimental design where the sensors were placed in the centre of the field.

@@

\section*{Proof}

@@

&=&\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top} 

@@

\begin{tabular} {c|c} 
	\hline\hline
	Symbol & Description \\ \hline
	$v(\mathbf{r},t)$ & mean membrane potential \\
	$g(\mathbf{r},t)$ & average action potential rate \\
	$\mathbf{r}$ & spatial location \\
	$t$ & time (s) \\
	$h(t)$ & post-synaptic response kernel \\
	$\eta(t)$ & Heaviside function \\
	$\zeta$ & inverse synaptic time constant \\
	$w(\mathbf{r},\mathbf{r}')$ & spatial connectivity kernel \\
	$f(\mathbf{r},t)$ & firing function rate \\
	$u(\mathbf{r},t)$ & external input \\
	$\Omega$ & spatial domain \\
	$f_{max}$ & maximal firing rate \\
	$\varsigma$ & slope of sigmoidal activation function \\
	$v_0$ & firing threshold \\
	$\delta(t)$ & Dirac-delta function \\
	$D$ & Temporal differential operator \\
	$L$ & number of basis function and dimension of state vector \\
	$e(\mathbf{r},t)$ & field disturbance, covariance $\gamma$\\
	$m(\mathbf{r},\mathbf{r}')$ & sensor kernel, variance $\sigma_m^2$ \\
	$\epsilon(\mathbf{r}_n,t)$ & observation noise, covariance $\Sigma_\epsilon$ \\
	$y(\mathbf{r}_n,t)$ & obseration \\
	$n$ & sensor index $n=1,..,N$ \\
	$T$ & time step \\
	$\xi$ & time constant parameter \\
	$\mathbf{\phi(r)}$ & vector of Gaussian basis functions \\
	$\mathbf{x}_t$ & state vector at time $t$ \\
	$\mathbf{\psi}$ & vector of connectivity kernel basis functions \\
	$\theta$ & vector of connectivity kernel parameters \\
	$\Gamma$ & inner product of field basis functions \\
	$q()$ & state function \\
	$k()$ & maps into to discrete state of input function \\
	$\mathbf{e}_t$ & state disturbance, covariance $\Sigma_e$ \\
	$\mathbf{C}$ & observation matrix \\
	$V(\nu)$ & Spectrum of the dynamic field \\ 
	$\nu$ & Spatial frequency \\
	$\nu_c$ & Spatial cut-off frequency \\
\end{tabular}
